---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# smoch.gp

<!-- badges: start -->
<!-- badges: end -->

Class imbalance is a common issue in medical datasets and compromises the performance of predictive models. Currently, over a hundred methods have been proposed to address this limitation, with the Synthetic Minority Over-sampling Technique  and Synthetic Minority Over-sampling Technique for Nominal and Continuous variables (SMOTE-NC)  being the most widely implemented over-sampling techniques in the medical domain. However, none of the methods developed to date take into account the heterogeneity of variables typically found in clinical data, which often results in the generation of synthetic observations without biological plausibility. The Synthetic Minority Over-sampling Convex Hull - Gower Podani (SMOCH-GP) method was proposed to address this gap in the literature. 

By employing the Gower distance with the Podani modification, the method ensures that the distances between observations are more reliable and reflective of real-world similarities. This approach also guarantees that the generation of new observations respects the nature of each variable type. For instance, if a variable is quantitative discrete, the resulting synthetic observations will also be integers rather than continuous values—a common issue observed in other over-sampling methods.


## Installation

You can install the development version of smoch.gp from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("beatrizlau1/smoch.gp")
```

## Example

This is a basic example which shows you how to apply to a imbalanced dataset: 

```{r example}
library(smoch.gp)
df <- data.frame(y=rep(as.factor(c('Yes', 'No')), times=c(90, 10)), x1=rnorm(100), x2=rnorm(100))
smoch.gp(y='y', x=c('x1','x2'), data=df, k=5, oversampling = 100, outlier = F)
```


